---
title: "RepData_PeerAssessment 1"
author: "Jacob Garcia"
date: "February 7, 2016"
output: html_document
---

#Reproducible Research: Peer Assessment 1
##Loading and preprocessing the data
* Set the working directory.
```{r}
setwd("C:/Users/Jacob/Desktop/RepData_PeerAssessment1")
getwd()
```
* As the needed data is in a zipped file, we'll need to unzip it first.
```{r, echo = TRUE} 
csvfile<-"activity.csv"
if(!exists(csvfile)){
     unzip(zipfile = "activity.zip")   
}
```
* Associate the data to the object Activity. Before this ,quickly review the data. You'll see that the "steps" column has NA values.
```{r, echo = TRUE} 
Activity<-read.csv(csvfile,header = T)
str(Activity)
```
* Transform the "date"" column to the correct Date format.
```{r, echo = TRUE} 
library(lubridate)
attach(Activity)
date<<-ymd(as.character(Activity$date))
Activity$time<-hm(paste(floor(interval/100),interval%%100,sep=" "))
detach(Activity)
str(Activity)
```

##Q1, What is mean total number of steps taken per day?
### Instructions
1. For this part of the assignment, you can ignore the missing values in the dataset.
2. Calculate the total number of steps taken per day
3. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
4. Calculate and report the mean and median of the total number of steps taken per day
```{r, echo = TRUE} 
StepsPerDay<-aggregate(steps~date,data=Activity,sum,na.rm=T)
# Notice result is different with the follow statement. 
#the day with all NA values are kept with 0 steps. 
#StepsPerDay<-aggregate(Activity$steps, by = list(date=Activity$date), FUN=sum, na.rm=T)
dim(StepsPerDay)

DailySteps<-StepsPerDay$steps
hist(x = DailySteps,breaks = 20)

options(digits = 2)
meanSteps<-mean(DailySteps)
medianSteps<-median(DailySteps)
```
* The mean of the total number of steps taken per day is : 10766, while the median is : 10765.

##Q2, What is the average daily activity pattern?
### Instructions
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo = TRUE} 
AverageIntervalSteps<-aggregate(steps~interval, data=Activity, mean, na.rm=T)
dim(AverageIntervalSteps)
```
```{r, echo = TRUE} 
with(AverageIntervalSteps,{
        plot(interval,steps,
             type="l",
             main = "Average number of steps taken",
             xlab="Interval, Time(hmm/hhmm)",
             ylab="average steps cross all days")
                
})
```
```{r, echo = TRUE} 
maxInterval<-AverageIntervalSteps[which.max(AverageIntervalSteps$steps),"interval"]
start<-paste(floor(maxInterval/100),maxInterval%%100,sep = ":")
end<-paste(floor(maxInterval/100),maxInterval%%100+5,sep = ":")
```
* The interval contains a maximun average steps of 835, from 8:35 to 8:40;
* This would be around the time a person would go to work.

## Q3, Imputing missing values
### Instructions
1. Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.
2. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
3. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
4. Create a new dataset that is equal to the original dataset but with the missing data filled in.
5. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
* Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo = TRUE} 
table(is.na(Activity))
```
```{r, echo = TRUE} 
sum(is.na(Activity$steps))
```
* Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
* Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo = TRUE} 
NewActivity<-Activity

for(i in 1:length(NewActivity$steps)){
        if(is.na(NewActivity$steps[i])){
                NewActivity$steps[i]<-
                        AverageIntervalSteps$steps[AverageIntervalSteps$interval==NewActivity$interval[i]]
        }
}

sum(is.na(NewActivity$steps))
```
```{r, echo = TRUE} 
sum(is.na(NewActivity$steps))
```
* Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r, echo = TRUE} 
StepsPerDay_New<-aggregate(steps~date,data=NewActivity,sum,na.rm=T)
DailySteps_New<-StepsPerDay_New$steps
opar<-par()
par(mfcol = c(1,2))
hist(DailySteps,20,col = "grey",ylim = c(0,15))
hist(DailySteps_New,20,col="blue",ylim =c(0,15))
```
```{r, echo = TRUE} 
par<-opar
options(digits = 2)
mean(DailySteps_New)
```
```{r, echo = TRUE} 
median(DailySteps_New)
```
* In the original data, if the data of a day is all NA, the day will be removed from the dataset. In the imputing histgram, the fix data with average interval steps are added into the most popular bucket.
```{r, echo = TRUE} 
with(AverageIntervalSteps,{
        plot(interval,steps,
             type="l",
             col="red",
             main = "Average number of steps taken",
             xlab="Interval, Time(hmm/hhmm)",
             ylab="average steps cross all days")
                
})
AverageIntervalStepsNew<-aggregate(steps~interval,data=NewActivity, mean, na.rm=T)
with(AverageIntervalStepsNew,{
        lines(x=interval,y=steps, col = "blue", lty ="twodash" )
})
legend("topright", 
       legend = c("Original","Fix Missing Value"), 
       col = c("red","blue"), 
       lty = c("solid","twodash"),cex=0.8)
```
```{r, echo = TRUE} 
maxInterval<-AverageIntervalStepsNew[which.max(AverageIntervalStepsNew$steps),"interval"]
start<-paste(floor(maxInterval/100),maxInterval%%100,sep = ":")
end<-paste(floor(maxInterval/100),maxInterval%%100+5,sep = ":")
```
* The interval contains a maximun average steps of 835, from 8:35 to 8:40;
* this is the same as the values estimated with missing values and shows that imputing missing data on the estimates of the total daily number of steps had no impact.
## Q4, Are there differences in activity patterns between weekdays and weekends?
### Instructions
1. For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.
2. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
3. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
* Firstly, plot the data of workdays and weekends together.
```{r, echo = TRUE} 
NewActivity$weekday<-wday(NewActivity$date,label=T)
NewActivity$DayType<-ifelse(NewActivity$weekday %in% c("Sun","Sat"),"weekends","workday")
AverageIntervalStepsWeekends<-
        aggregate(NewActivity$steps,
                  by=list(DayType=NewActivity$DayType,interval=NewActivity$interval),
                  FUN = mean,
                  na.rm=T)
library(ggplot2)
qplot(data=AverageIntervalStepsWeekends,
      x = interval,
      y = x,
      geom="line",
      col=DayType,
      xlab="Interval, Time(hmm/hhmm)",
      ylab="average steps cross all days")+theme(legend.position = "top")
```

* Secondly, calculate the compare the mean/median between weekdays and weekends.
```{r, echo = TRUE} 
attach(AverageIntervalStepsWeekends)
aggregate(x,FUN = mean,by=list(dayType=DayType))
```
```{r, echo = TRUE} 
aggregate(x,FUN = median,by=list(dayType=DayType))
```
```{r, echo = TRUE} 
detach(AverageIntervalStepsWeekends)
```
* We can see that the average steps of the weekends is more than weekdays;
And the time series pattern is also different:
a. In the morning(5~10 o' clock), weekend < workdays;
b. In the daylight(10~20 o' clock), weekend > workdays;

* Finally, plot the data by weekdays to further understand.
```{r, echo = TRUE} 
AverageIntervalStepsPerWeekday<-
        aggregate(NewActivity$steps,
                  by=list(weekday=NewActivity$weekday,interval=NewActivity$interval),
                  FUN = mean,
                  na.rm=T)

qplot(data=AverageIntervalStepsPerWeekday,
      x = interval,
      y = x, 
      geom="line",
      facets =.~ weekday, 
      ylab="average steps",
      xlab="Interval, Time(hmm/hhmm)")
```

* Finished